\documentclass[11pt,a5paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage[margin=0.47in]{geometry}


\newtheorem{theorem}{Example}
\newtheorem{exercise}{Exercise}
\newtheorem*{Theorem}{Theorem}

\title{\textbf{Mean Inequalities}}
\date{Week 5}
\author{Miroslav Stankovic\\ Marko Puza}
\begin{document}
\maketitle

\section{Theory}
\subsection*{Basics}

Inequality problems are a very broad topic. In this text, we will 
mostly focus the inequalities between means and how to apply them 
to solve wide range of inequalities. \\

\noindent We will be only concerned with real numbers (often only 
non-negative real numbers). One of basic properties of real numbers 
is that the square of a real number is non-negative, thus so is sum 
of squares. This is the most basic useful inequality.

\begin{Theorem} \emph{Sum of squares inequality.}
	If $x_1, x_2, \dots x_n \in \mathbf{R}$ and 
	$\alpha_1, \alpha_2, \dots \alpha_n > 0$ then 
	$\sum_{i=1}^n \alpha_1x_1^2\ge0$ with equality if and only if 
	$x_1 = x_2 = \dots = x_n = 0$.
\end{Theorem}

\begin{theorem} Prove that $x^2 + y^2 + 2 \ge 2x + 2y$ and find when 
there is equality.
\end{theorem}
\begin{proof} 
	\begin{align*} 
		x^2 + y^2 + 2 & \ge 2x + 2y \\
		x^2 -2x +1 + y^2 -2y + 1 & \ge 0 \\
		(x-1)^2 + (y-1)^2 & \ge 0
	\end{align*}
	So our inequality is equivalent to the $(x-1)^2 + (y-1)^2 \ge 0$, 
	which is Sum of squares inequality with equality for $x=y=1$
\end{proof}
\begin{exercise} Prove that $4x^2 + y^2 + 3 \ge 4x + 2y$.
\end{exercise}

\noindent Thanks to the work of Emil Artin and Charles Delzell 
it is now known, that it is possible to prove any inequality 
involving rational functions simply by reducing it to a sum of 
squares. However, to do so would be outstandingly tedious and 
so we introduce much more powerful method - Arithmetic-Geometric 
mean inequality.

\subsection*{Arithmetic-Geometric mean inequality}

A well known inequality, often used in problem solving, is an 
inequality between arithmetic and geometric mean. It is often 
called AM-GM inequality or simply AG.

\begin{Theorem} \emph{AG.}
For nonegative real numbers $x_1, x_2, \dots x_n$ : 
\[\sqrt[n]{x_1x_2\dots x_n} \le  \frac{x_1 + x_2 + \dots + x_n}{n}\] 
with equality if and only if $x_1 = x_2 = \dots = x_n$. \\
\end{Theorem}

\noindent We will prove AG for $n=2$. The AG becomes 
$\sqrt{x_1x_2} \le  \frac{x_1 + x_2}{2}$. 
\begin{proof} Using that square of a real is non-negative and basic 
algebraic operations:
	\begin{align*}
	0 & \le (x-y)^2 \\
	 & = x^2 - 2xy + y^2 \\
	 & = x^2 + 2xy - y^2 -4xy \\
	 & = (x+y)^2 - 4xy 
	\end{align*}
	Thus $4xy \le (x+y)^2$ from which 
	$\sqrt{x_1x_2} \le  \frac{x_1 + x_2}{2}$ follows imidiately. 
\end{proof}

\begin{theorem}
	Prove that for possitive real numbers $x, y, z$: 
	$x^3 + y^3 + z^3 \ge x^2y + y^2z + z^2x$.
\end{theorem}

\begin{proof} Let's look at each expression on the right-hand side 
separatelly. We can treat $x^2y$ as product of three numbers 
$x\cdot x\cdot y$ and using AG we have 
\[x^2y = x\cdot x\cdot y = \sqrt[3]{x^3x^3y^3} \le\frac{x^3 + x^3 + y^3}{3}\]
We get very similar inequalities for $y^2z$ and $z^2x$ - summing 
those three up we get exactly the inequality we wanted to prove.
\end{proof}

\begin{exercise} Prove that for possitive real numbers $x, y, z$: 
$x^4 + y^4 + z^4 \ge x^2yz + xy^2z + xyz^2$.
\end{exercise}

\subsection*{Weights and Power Mean Inequality}

There are some generalizations of AM-GM inequality, first of them uses 
so called weighted means. 
\begin{Theorem}\emph{Weighted AG.} 
	For nonegative real numbers $x_1, x_2, \dots x_n$ and non-negative real 
	weights $w_1, w_2, \dots w_n$ where $w = w_1 + w_2 + \dots + w_n > 0$:
\[\sqrt[w]{x_1^{w_1}x_2^{w_2}\dots x_n^{w_n}} \le  \frac{w_1x_1 + w_2x_2 + \dots + w_nx_n}{w}\] 
	with equality if and only if all the $x_k$ with $w_k > 0$ are equal. \\
\end{Theorem}

\begin{theorem}
	Prove that for possitive real $x, y, z$: 
	$\frac{x^2}{2} + \frac{y^3}{3} + \frac{z^6}{6} \ge xyz$.
\end{theorem}

\begin{proof} 
	Notice that $\frac{1}{2} + \frac{1}{3} + \frac{1}{6} = 1$ Using weights $\frac{1}{2}$, $\frac{1}{3}$, and $\frac{1}{6}$, the weighted AG for numbers $x^2$, $y^3$, and $z^6$ gives us exactly the inequality we wanted to prove.
\end{proof}

\begin{exercise} Prove that for possitive real $x, y, z$: 
	$\frac{1}{2x^2} + \frac{1}{3y^3} + \frac{1}{6z^6} \ge \frac{1}{xyz}$.
\end{exercise}

Another very powerful generalization arises when looking at other means, 
not just arithmetic or geometric. The generalized mean, or power mean, 
is defined as follows:

\begin{Theorem} \emph{AG.}
If $p$ is a non-zero real number and $x_1, x_2, \dots , x_n$ are positive 
integers then power mean with exponent $p$ is
\[M_p(x_1, x_2, \dots, x_n) 
= \bigg(\dfrac{x_1^p + x_2^p + \dots + x_n^p}{n}\bigg)^\frac{1}{p}\]
\end{Theorem}

\noindent We can see, that for $M_1$ is arithmetic mean. For $p=0$, 
the $M_p$ is not defined, however it can be shown that 
\[\lim_{p\to 0} M_p(x_1, x_2, \dots, x_n) = \sqrt[n]{x_1x_2\dots x_n}\]
The geometric mean is therefor assigned as $M_0$. So far we have discussed 
$M_0 \le M_1$. So how do other means relate to this?

\begin{Theorem} \emph{Generalized mean inequality.}
For any positive real numbers $x_1, \dots,  x_n$ if $p < q$ then $M_p(x_1, \dots, x_n) < M_q(x_1, \dots, x_n)$.

\end{Theorem}

\section{Problems}

\begin{enumerate}
	\subsection*{Easy}
	\item{Demonstrate that if $a_1a_2\cdots a_n=1$ then 
	$a_1+a_2+\cdots +a_n\ge n$.}

	\subsection*{Medium}
	
	\subsection*{Difficult}

	\subsection*{Extra}	
	\item{Prove AG using induction.}
	
	\item{Prove that power mean tends to geometric mean as the power 
	tends to zero.}
\end{enumerate}

\begin{thebibliography}{9}
\bibitem{KMS} Ondrej Budáč, Tomáš Jurík, and Ján Mazák. 
	\emph{Zbierka úloh KMS}. Trojsten, Bratislava, 2010.

\end{thebibliography}

\section{Hints}
\begin{enumerate}
	\subsection*{Easy}

	\subsection*{Medium}
	
	\subsection*{Difficult}

	\subsection*{Extra}	
	\item{Use the three step induction. Show that if AG holds for $n$ 
	numbers then it also holds for $2n$ numbers and then that if it 
	holds for $m$ numbers it also holds for $2m-1$ numbers. (Why this 
	induction works?)}
	
	\item{Use the L'Hôpital's rule.}
\end{enumerate}

\end{document}

